Testing atlantisom: true biomass output from Atlantis

### 2019-04-29 TrueBioTest.Rmd
add N based survey and plotting

```{r message=FALSE, warning=FALSE}
library(tidyr)
require(dplyr)
library(ggplot2)
library(data.table)
library(here)
```

@@

```{r load_Rdata, message=FALSE, warning=FALSE, eval=FALSE}

@@

```{r surveyNbased , eval = FALSE}

@@

@@

>>>>

```{r message=FALSE, warning=FALSE}
library(tidyr)
require(dplyr)
library(ggplot2)
library(data.table)
library(here)
library(ggforce)
library(ggthemes)
```

@@

```{r load_Rdata, message=FALSE, warning=FALSE}

@@

Comparing our (census) survey based on true biomass from above with the Atlantis output file "[modelscenario]BiomIndx.txt" should give us a perfect match. Note that the our (census) survey may have more sampling in time than the Atlantis output file.

```{r matchB, fig.cap="Testing whether the survey census gives the same results as the Atlantis output biomass index file; first 9 species."}
# plot some comparisons with Atlantis output

# read Atlantis output files
if(initCCA) {
  atBtxt2 <- read.table(here("atlantisoutput","CalCurrentSummitScenario1","outputCCV3BiomIndx.txt"), header=T)
  groupslookup <- load_fgs(dir = d.name, functional.groups.file)
  surveyB_frombio <- CCAsurveyB_frombio
}

if(initNEUS) {
  atBtxt2 <- read.table(here("atlantisoutput","NEUStest20160303","neusDynEffort_Test1_BiomIndx.txt"), header=T)
  groupslookup <- load_fgs(dir = d.name, functional.groups.file)
  surveyB_frombio <- NEUSsurveyB_frombio
}

if(initNOBA) {
  atBtxt2 <- read.table(here("atlantisoutput","NOBACERESGlobalSustainability","outputnordic_runresults_01BiomIndx.txt"), header=T) 
  groupslookup <- load_fgs(dir = d.name, functional.groups.file)
  surveyB_frombio <- NOBAsurveyB_frombio
}

# lookup the matching names, put in time, species, biomass column format
# WARNING hardcoded for output with last species group as DIN
groupslookup <- groupslookup %>%
  filter(IsTurnedOn > 0)

atBtxt2tidy <- atBtxt2 %>%
  select(Time:DIN) %>%
  #select(Time, FPL:DIN) %>%
  rename_(.dots=with(groupslookup, setNames(as.list(as.character(Code)), Name))) %>%
  gather(species, biomass, -Time) %>%
  filter(species == levels(surveyB_frombio$species))

#all species comparison, time intervals hardcoded for NEUS and NOBA
compareB <-ggplot() +
  geom_line(data=surveyB_frombio, aes(x=time/5,y=atoutput, color="survey census B"), 
            alpha = 3/10) +
  geom_point(data=atBtxt2tidy, aes(x=Time/365,y=biomass, color="txt output true B"),
             alpha = 3/10) + 
  theme_tufte() +
  theme(legend.position = "top") +
  labs(colour=scenario.name)

compareB + 
  facet_wrap_paginate(~species, ncol=3, nrow = 3, page = 1, scales="free")

```{r matchBp2, fig.cap="Testing whether the survey census gives the same results as the Atlantis output biomass index file; next 9 species."}
compareB + 
  facet_wrap_paginate(~species, ncol=3, nrow = 3, page = 2, scales="free")

```{r matchBp3, fig.cap="Testing whether the survey census gives the same results as the Atlantis output biomass index file; last 9 species. Depending on the model, we may not show all species in this example."}
compareB + 
  facet_wrap_paginate(~species, ncol=3, nrow = 3, page = 3, scales="free")

After comparing the survey census based on the biomass output of `run_truth`, we now look at a biomass index that is estimated based on numbers in the survey census and an average weight at age. This is more complex, but may be closer to the way some assessments handle survey observations.

```{r surveyNbased, fig.cap="Check weight at age generated from biomass_age$atoutput divided by nums$atoutput for first 32 groups.", fig.show="hold"}

@@

# as above, make up a constant 0 cv for testing
surv_cv <- data.frame(species=survspp, cv=rep(0.0,length(survspp)))

# now we need a weight at age to convert numbers to weight
# if I do biomass_age$atoutput divided by nums$atoutput I should have wt@age, yes?

nums_test <- with(datN, aggregate(atoutput, list(species, agecl), sum))
names(nums_test) <- c("species", "agecl", "nums")

bio_test  <- with(datB, aggregate(atoutput, list(species, agecl), sum))
names(bio_test) <- c("species", "agecl", "wt")

calcwtage <- merge(bio_test, nums_test) %>%
  mutate(wtAtAge=wt/nums)

wtagecheck <- ggplot(calcwtage, aes(x=agecl, y=wtAtAge)) +
  geom_point() + 
  theme_tufte() +
  theme(legend.position = "top") +
  labs(colour=scenario.name)

wtagecheck + 
  facet_wrap_paginate(~species, ncol=4, nrow=4, page=1, scales="free")

wtagecheck + 
  facet_wrap_paginate(~species, ncol=4, nrow=4, page=2, scales="free")

# variable name needs to be wtAtAge or doesnt work
wtage <- calcwtage %>%
  select(species, agecl, wtAtAge)

surveyB_fromN <- sample_survey_biomass(survey_testNall, surv_cv, wtage)

if(initCCA) CCAsurveyB_fromN <- surveyB_fromN
if(initNEUS) NEUSsurveyB_froN <- surveyB_fromN
if(initNOBA) NOBAsurveyB_fromN <- surveyB_fromN

@@

Now we plot the results of the numbers-based (census) survey biomass against the same Atlantis output file "[modelscenario]BiomIndx.txt" as above.

```{r matchBfromN, fig.cap="Testing whether the survey census based on numbers converted to biomass gives the same results as the Atlantis output biomass index file; first 9 species." }

if(initCCA) surveyB_fromN <- CCAsurveyB_fromN
if(initNEUS) surveyB_froN <- NEUSsurveyB_fromN
if(initNOBA) surveyB_fromN <- NOBAsurveyB_fromN

#all species comparison, time intervals hardcoded for NEUS and NOBA
compareB_N <-ggplot() +
  geom_line(data=surveyB_fromN, aes(x=time/5,y=atoutput, color="survey census N->B"), 
            alpha = 3/10) +
  geom_point(data=atBtxt2tidy, aes(x=Time/365,y=biomass, color="txt output true B"),
             alpha = 3/10) + 
  theme_tufte() +
  theme(legend.position = "top") +
  labs(colour=scenario.name)

compareB_N + 
  facet_wrap_paginate(~species, ncol=3, nrow = 3, page = 1, scales="free")

```{r matchBfromN2, fig.cap="Testing whether the survey census based on numbers converted to biomass gives the same results as the Atlantis output biomass index file; next 9 species."}

compareB_N + 
  facet_wrap_paginate(~species, ncol=3, nrow = 3, page = 2, scales="free")

```{r matchBfromN3, fig.cap="Testing whether the survey census based on numbers converted to biomass gives the same results as the Atlantis output biomass index file; last 9 species.Depending on the model, we may not show all species in this example."}

compareB_N + 
  facet_wrap_paginate(~species, ncol=3, nrow = 3, page = 3, scales="free")
